---
title: "Pollination and frugivory in ivy: Electronic R code"
author: "Oi Ki Miranda Lam"
date: "2023-12-13"
output:
  html_document:
    toc: TRUE
---

## Pollinator visits

```{r message = FALSE, results = FALSE, warning = FALSE}
# Load required packages
library(readxl)
library(ggplot2)
library(ggpubr)
```


```{r}
# Import and format data
survey_data <- data.frame(read.csv("Results from pollinator surveys.csv"))
df_count <- survey_data[c('Site','No..of.umbels','Wasps','Total.no..of.visits')]
df_count$Site <- factor(df_count$Site, levels = c("TM", "BG"))
```

#### Generalised Linear Model for number of visits by all insects
```{r}
model1 <- glm(Total.no..of.visits ~ Site + offset(log(No..of.umbels)), family = poisson(link = 'log'), data =df_count)
summary(model1)

#effect_plot(model1, pred = Site,interval = TRUE, data = df_count, partial.residuals = TRUE, jitter = .1)
```

##### Model validation
```{r}
plot(model1)
theta <-model1$deviance/model1$df.residual
theta

devresid <- resid(model1, type = 'deviance')
plot(devresid ~ model1$fitted.values)
plot(devresid ~ df_count$Site)

```

#### Generalised Linear Model for number of visits by wasps only
```{r}
# hist(df_count$Wasps)
wasp1 <- glm(Wasps ~ Site + offset(log(No..of.umbels)), family = quasipoisson (link = 'log'), data = df_count)
summary(wasp1)

```

##### Model validation
```{r}
plot(wasp1)
theta <-wasp1$deviance/wasp1$df.residual
theta

devresid <- resid(model1, type = 'deviance')
plot(devresid ~ model1$fitted.values)
plot(devresid ~ df_count$Site)

```

## Pollinator community

```{r}
# Import & format data

grp_counts <- as.table(rbind(c( sum(survey_data$Bumblebees[survey_data$Site == 'TM']),
                                  sum(survey_data$Honey.bees[survey_data$Site == 'TM']),
                                  sum(survey_data$Solitary.bees[survey_data$Site == 'TM']),
                                  sum(survey_data$Wasps[survey_data$Site == 'TM']),
                                  sum(survey_data$Hoverflies[survey_data$Site == 'TM']),
                                  sum(survey_data$Other.flies[survey_data$Site == 'TM']),
                                  sum(survey_data$Butterflies.and.Moths[survey_data$Site == 'TM']),
                                  sum(survey_data$Small.insects...3mm.[survey_data$Site == 'TM'])
                                  ),
                               c( sum(survey_data$Bumblebees[survey_data$Site == 'BG']),
                                  sum(survey_data$Honey.bees[survey_data$Site == 'BG']),
                                  sum(survey_data$Solitary.bees[survey_data$Site == 'BG']),
                                  sum(survey_data$Wasps[survey_data$Site == 'BG']),
                                  sum(survey_data$Hoverflies[survey_data$Site == 'BG']),
                                  sum(survey_data$Other.flies[survey_data$Site == 'BG']),
                                  sum(survey_data$Butterflies.and.Moths[survey_data$Site == 'BG']),
                                  sum(survey_data$Small.insects...3mm.[survey_data$Site == 'BG'])
                                 )))
                                  
dimnames(grp_counts) <- list(site = c("TM", "BG"),
                               groups = c("Bumblebee",
                                          "Honey bee",
                                          "Solitary bee",
                                          "Wasp",
                                          "Hoverfly",
                                          "Other flies",
                                          "Butterflies & Moths",
                                          "Small insects (<3mm)")) 
# Display data as table
knitr::kable(grp_counts, caption = "No. of pollinators recorded" )

```

#### Fisher's exact test
```{r}
fisher.test(grp_counts, simulate.p.value = TRUE)
```

## Generalised Linear Model for fruiting success

```{r}
# Import & format data
df_frset <- read.csv("Fruit set.csv")
df_frset$Site <- factor(df_frset$Site, levels = c("TM", "BG"))
df_frset$Treatment <- factor(df_frset$Treatment, levels = c("Bagged", "Open", "Hand Pollinated"))

df_frset$fruit_set <- cbind(df_frset$Success, df_frset$Failure) # create success/fail for fruit set
```

#### Quasibinomial fruit set stepwise backwards elimination

Full saturated model:

```{r}
setmod1 <- glm(fruit_set ~ Site + Treatment + Site:Treatment, data=df_frset, family=quasibinomial(link="logit"))
summary(setmod1)
# Check which variable to drop next
drop1(setmod1, test="F")
```

Interaction variable (Site:Treatment) is dropped in the next model.

```{r}
setmod2 <- glm(fruit_set ~ Site + Treatment, data=df_frset, family=quasibinomial(link="logit")) 
anova(setmod1, setmod2, test = "F") #Comparing fit of models, retain term if fit is significantly different
drop1(setmod2, test="F")
```

The 'Site' variable will be dropped in the next model.

```{r}
setmod3 <- glm(fruit_set ~ Treatment, data=df_frset,
family=quasibinomial (link="logit"))
anova(setmod2, setmod3, test = "F")
drop1(setmod3, test="F")
```

Results suggest that dropping the 'Treatment' variable will lead to significant differences in model fit. The variable was dropped in the next model to see if this is true.

```{r}
setmod4 <- glm(fruit_set ~ 1, data=df_frset, family=quasibinomial)
anova(setmod3, setmod4, test = "F")

```

Dropping the 'Treatment' variable led to significance difference in model fit, therefore this variable is retained in the final model.

#### Model validation

##### Diagnostic plots

```{r}
plot(setmod3) 
```

#### Post-hoc tukey HSD test

```{r message = FALSE, results = FALSE, warning = FALSE}
# Load required packages
library(multcomp)
```

```{r}
summary(glht(setmod3, mcp(Treatment="Tukey")))
```
#### Describe data

```{r message = FALSE, warning = FALSE}
df_frset %>% group_by(Treatment) %>% dplyr::summarise(mean(Percentage),
                                                      sd = sd(Percentage),
                                                      n = length(Percentage),
                                                      se = sd / sqrt(n)
                                                      )
```

## Generalised Linear Model for fruit size

```{r}
# Import & format data
dfsize <- read.csv("Fruit size.csv")
dfsize$Site <- factor(dfsize$Site,levels = c("TM", "BG")) 
dfsize$Treatment <- factor(dfsize$Treatment, levels = c("Bagged", "Open", "Hand pollinated"))
dfsize_mean <- dfsize %>% group_by(Site, Treatment, Replicate.no.) %>% dplyr::summarise(Size = mean(Size..mm.))

# dfsize <- read_excel("Ivy pollination study supporting data 131223.xlsx", sheet = "Fruit size R")
# dfsize$Site <- factor(dfsize$Site,levels = c("TM", "BG")) 
# dfsize$Treatment <- factor(dfsize$Treatment, levels = c("Bagged", "Open", "Hand Pollinated"))
```

#### Stepwise backwards elimination

Full saturated model:

```{r}
sizemod1 <- glm(Size ~ Site + Treatment + Site:Treatment, data=dfsize_mean, family=gaussian (link="identity"))
summary (sizemod1)
drop1(sizemod1, test="F") # Use to check variables to drop
```

This shows that the interaction variable (Site:Treatment) can be dropped without leading to significant differences in model fit.

```{r}
sizemod2 <- glm(Size ~ Site + Treatment, data=dfsize_mean, family=gaussian (link="identity"))
anova(sizemod1, sizemod2, test = "F")
drop1(sizemod2, test="F")
```

This shows that the 'Site' variable can be dropped.

```{r}
sizemod3 <- glm(Size ~ Treatment, data=dfsize_mean, family=gaussian (link="identity"))
anova(sizemod2, sizemod3, test = "F")
drop1(sizemod3, test="F")
```

This shows that dropping the 'Treatment' variable will lead to significant differences in model fit. This was tested in the next model.

```{r}
sizemod4 <- glm(Size ~ 1, data=dfsize_mean, family=gaussian (link="identity"))
anova (sizemod3, sizemod4, test = "F")
```

Dropping the 'Treatment' variable led to significance difference in model fit, therefore this variable is retained in the final model.

#### Model validation

##### Diagnostic plots

```{r}
plot(sizemod3) 
```

#### Post-hoc tukey HSD test

```{r message = FALSE, results = FALSE, warning = FALSE}
# Load required packages
library(multcomp)
```

```{r}
summary(glht(sizemod3, mcp(Treatment="Tukey")))
```

#### Describe data
```{r}
dfsize %>% group_by(Treatment) %>% dplyr::summarise(mean(Size..mm.),
                                                      sd = sd(Size..mm.),
                                                      n = length(Size..mm.),
                                                      se = sd / sqrt(n)
                                                      )
```

## Survival analysis

```{r message = FALSE, results = FALSE, warning = FALSE}
# Load required packages
library(survival)
library(survminer)
library(dplyr) 
library(ggpubr) 

# Import & format data
survdata <- read.csv("Survival analysis.csv")
TM <- subset(survdata, Site == "TM") 
BG <- subset(survdata, Site == "BG")
```

#### Trumpington Meadows:

##### Fit survival data using the Kaplan-Meier method

```{r}
surv_object_TM <- Surv(time = TM$Survival.time, event = TM$Censored) 
TMfit1 <- survfit(surv_object_TM ~ Treatment, data = TM) 
summary(TMfit1, times = c(4, 14))
```

##### Log-rank test

```{r}
survdiff(Surv(time = TM$Survival.time, event = TM$Censored) ~ Treatment, data = TM)

```

##### Plotting survival curve

```{r}
ggsurvplot(TMfit1, data = TM, 
           conf.int = TRUE, 
           legend.title = "Treatment", 
           legend.labs = c("Bagged","Hand Pollinated", "Open"), 
           font.x = 17, font.y = 17, font.legend = 17, 
           legend = "right", 
           palette = c("Blue", "azure4", "Red"), pval = TRUE)
```

#### Botanic Garden:

##### Fit survival data using the Kaplan-Meier method

```{r}
surv_object_BG <- Surv(time = BG$Survival.time, event = BG$Censored) 
BGfit1 <- survfit(surv_object_BG ~ Treatment, data = BG)
summary(BGfit1, times = c(4, 14)) 
```

##### Log-rank test

```{r}
survdiff(Surv(time = BG$Survival.time, event = BG$Censored) ~ Treatment, data = BG)
```

##### Plotting survival curve

```{r}
ggsurvplot(BGfit1, data = BG, 
           conf.int = TRUE,
           legend.title = "Treatment", 
           legend.labs = c("Bagged","Hand Pollinated", "Open"), 
           font.x = 17, font.y = 17, font.legend = 17,
           legend = "right",
           palette = c("Blue", "azure4", "Red"), pval = TRUE)
```

## Appendix - Species-accumulation curves and sampling completedness (Fig. S1)

```{r}
# Load required packages
library(iNEXT)
library(ggplot2)

# Import & format data
spacc <- read.csv("iNEXT.csv")
spacc_df <- data.frame(spacc$Site, spacc$Species, spacc$Quantity)
spacc_num <- list(TM=as.numeric(spacc_df[which(spacc_df$spacc.Site==1),3]),
                 BG=as.numeric(spacc_df[which(spacc_df$spacc.Site==2),3]))
```

#### Plotting species-accumulation curves

```{r warning = FALSE, message = FALSE}
new.curve <- iNEXT(spacc_num, q=0, datatype="abundance", endpoint = 500)
new.curve
DataInfo(spacc_num, datatype="abundance")

ggiNEXT(new.curve, type=1) + theme_bw(base_size = 18) + scale_colour_manual(values=c("red", "blue")) + scale_fill_manual(values=c("red", "blue"))
```

#### Species richness for each site

```{r warning = FALSE, message = FALSE}
ChaoRichness(as.numeric(spacc_df[which(spacc_df$spacc.Site==1),3])) #TM richness
ChaoRichness(as.numeric(spacc_df[which(spacc_df$spacc.Site==2),3])) #BG richness
```

#### Plotting sample completedness curves

```{r warning=FALSE, message = FALSE}
ggiNEXT(new.curve, type=2) + theme_bw(base_size = 18) + scale_colour_manual(values=c("red", "blue")) + scale_fill_manual(values=c("red", "blue"))
```
